Qualifying Report

### 2019-01-29 Source/QualifyingReport/QualifyingReport.tex
more edits

@@

Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task). There are a number of benefits to using single word tasks. From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well. However, this approach does not take into account the symbiotic nature of language processes.

@@

Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity. According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD. If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD. An example of a semantic complexity measure is the concept of idea density. Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}. Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic (described below) which are used to measure the lexical diversity of a given piece of text and/or utterance.  This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.

@@

\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
\begin{equation} \label{x3}
W = N^{V(-0.165)}
where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject. Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers. \newline
\subsubsection{Honore's Statistic (R)}
Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation:
\begin{equation} \label{x4}
%% Check to see if this is accurate.
R = (100 \log N) / (1 - V1/V)
where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.

@@

Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}. Bschor et al. (2001) examined Cookie Theft picture descriptions at four stages of AD. They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls.

@@

Conciseness has been defined as the number of words a speaker uses to express ideas. The theory is that people with AD will need more words to convey ideas because of word-finding related behaviours such as circumlocutions and repetitions. Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index. Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}. Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life.

@@

Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD. Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words. In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.

@@

Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms. Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity. Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences. Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD. The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.

@@

\subsubsection{Linguistic Inquiry and Word Count}
The premise behind the Linguistic Inquiry and Word Count (LIWC) is the words we speak inform an observer of the state of mind of an individual. To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised. 
% Al-Mosaiwi - In an Absolute State: Elevated Use of Absolutist Words Is a Marker Specific to Anxiety, Depression, and Suicidal Ideation 
A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common. 
% Sonnenschein - Linguistic analysis of patients with mood and anxiety disorders during cognitive behavioral therapy
The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety. 
\par
A framework like this could be used as another tool to explore themes within a persons speech.

@@

There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease. This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2008}. Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech. An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W. Bush who acted as a matched control based on Age \cite{Berisha2015}. They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia. These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}. 
/par   

@@

\section{Discussion}
\section{Conclusion}

@@

\section{Ethical Considerations}
In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations. I am aware of the code of human  research ethics published by the British Psychological Society \cite{Society2012} and my research has been designed in a way that fits with these guidelines.
\subsection{Risk}
I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition. However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependant or unequal relationship. Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
\subsection{Informed Consent}
Consent will be sought overtly at the recruitment and initial data collection point by way of an information sheet and associated consent form. At other data collection points, I will have information sheets available if participants request them and if such a request is made then I will seek consent again at this stage. As part of this process, I will be supplying my participants with my contact details should they have any question or for the purposes of withdrawing consent. These consent forms will be stored in a secure, locked cabinet at the university. 
\subsection{Confidentiality}
Given the nature of the study and the possibility of disclosures of suicidal ideation, it is important that confidentiality is discussed at each data collection point. I will also provide an information sheet which details the confidentiality statement and also the situations in which I may need to break confidentiality. In addition, I will make a verbal confidentiality statement at each collection point and seek acknowledgement of their understanding and permission to continue. The spoken confidentiality statement will be as follows.
\begin{center}
	\textit{"Anything you say to me today will be confidential unless I feel that there is a risk to you, or to someone else. If this is the case, I may need to speak to or write to third parties such as your GP. I will endeavor to discuss this communications with yourself where it is possible to do so. Is this ok?"}
\end{center}
\subsection{Disclosure of suicidal ideation or other risks}
As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9). This is something to explore further. As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.  Optionally with the tape recorder turned on, or off as requested.  Any text recorded during the risk assessment will not be included in the language analysis. Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided. A discussion will then be made about whether it is appropriate to notify the GP of the disclosure. In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact. (Crisis Team and GP). In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.

>>>>

There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem. This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention. 

@@

Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task). There are a number of benefits to using single word tasks. From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well. However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks. More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities. They also do not require the logical and efficient organisation of ideas. 

@@

Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity. According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD. If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD. An example of a semantic complexity measure is the concept of idea density. Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}. Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic (described below) which are used to measure the lexical diversity of a given piece of text and/or utterance.  This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.

@@

@@

Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}. Bschor et al. (2001) examined Cookie Theft picture descriptions at four stages of AD. They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls \cite{Bschor2001}.

@@

Conciseness has been defined as the number of words a speaker uses to express ideas. The theory is that people with AD will need more words to convey ideas because of word-finding related behaviours such as circumlocutions and repetitions. Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index. Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}. Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.

@@

Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD. Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}. In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.

@@

Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms. Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity \cite{Orimaye2017}. Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences. Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD. The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.

@@

\subsubsection{Linguistic Inquiry and Word Count}
The premise behind the Linguistic Inquiry and Word Count (LIWC) is the words we speak inform an observer of the state of mind of an individual. To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised \cite{Pennebaker2015}. 
% Al-Mosaiwi - In an Absolute State: Elevated Use of Absolutist Words Is a Marker Specific to Anxiety, Depression, and Suicidal Ideation 
A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common. 
% Sonnenschein - Linguistic analysis of patients with mood and anxiety disorders during cognitive behavioral therapy
The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety. 
\par
A framework like this could be used as another tool to explore themes within a persons speech.

@@

There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease. This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}. Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech. An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W. Bush who acted as a matched control based on Age \cite{Berisha2015}. They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia. These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}. 

@@

	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
		& RR & GHWB & DJT \\ \hline
		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline

	\end{tabular}
	\caption{\label{tab:table-name}NLTK tasks and functionality}
	\end{center} 
\end{table}
\section{Discussion}
\section{Conclusion}

@@

\section{Ethical Considerations}
In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations. I am aware of the code of human  research ethics published by the British Psychological Society \cite{BritishPsychologicalSociety2014} and my research has been designed in a way that fits with these guidelines.
\subsection{Risk}
I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition. However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependant or unequal relationship. Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
\subsection{Informed Consent}
Consent will be sought overtly at the recruitment and initial data collection point by way of an information sheet and associated consent form. At other data collection points, I will have information sheets available if participants request them and if such a request is made then I will seek consent again at this stage. As part of this process, I will be supplying my participants with my contact details should they have any question or for the purposes of withdrawing consent. These consent forms will be stored in a secure, locked cabinet at the university. 
\subsection{Confidentiality}
Given the nature of the study and the possibility of disclosures of suicidal ideation, it is important that confidentiality is discussed at each data collection point. I will also provide an information sheet which details the confidentiality statement and also the situations in which I may need to break confidentiality. In addition, I will make a verbal confidentiality statement at each collection point and seek acknowledgement of their understanding and permission to continue. The spoken confidentiality statement will be as follows.
\begin{center}
	\textit{"Anything you say to me today will be confidential unless I feel that there is a risk to you, or to someone else. If this is the case, I may need to speak to or write to third parties such as your GP. I will endeavor to discuss this communications with yourself where it is possible to do so. Is this ok?"}
\end{center}
\subsection{Disclosure of suicidal ideation or other risks}
As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9). This is something to explore further. As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.  Optionally with the tape recorder turned on, or off as requested.  Any text recorded during the risk assessment will not be included in the language analysis. Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided. A discussion will then be made about whether it is appropriate to notify the GP of the disclosure. In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact. (Crisis Team and GP). In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
